---
title: "Clustering with Credit Card Customers"
output:
  pdf_document: default
  html_notebook: default
---

```{r q1}
library(tidyverse) 
library(tidymodels)     
library(DataExplorer) 
library(tidyclust)    
library(dbscan)       
library(factoextra)
```

# Load Data

```{r q2}
df <- read_csv("credit_card_customers.csv", show_col_types = FALSE)
```

# Data Cleaning

```{r q3}
df <- df %>%
  mutate(
    CLIENTNUM = as.character(CLIENTNUM),
    Attrition_Flag = recode(Attrition_Flag,
                            "Existing Customer" = "Retained",
                            "Attrited Customer" = "Churned"),
    Gender = factor(Gender, ordered = FALSE),
    Marital_Status = factor(Marital_Status, ordered = FALSE),
    Education_Level = as.integer(factor(Education_Level, levels = c("Unknown", "Uneducated", "High School",
                                                         "College", "Graduate", "Post-Graduate",
                                                         "Doctorate"), ordered = TRUE)),
    Income_Category = as.integer(factor(Income_Category, levels = c("Unknown", "Less than $40K", 
                                                          "$40K - $60K", "$60K - $80K", 
                                                          "$80K - $120K", "$120K +"), ordered = TRUE)),
    Card_Category = as.integer(factor(Card_Category, c("Blue", "Silver", "Gold", "Platinum"), ordered = TRUE))
  )
```

# Data Exploration

```{r q4}
plot_intro(df)
```

## Bar plots of categorical variables

```{r q5}
plot_bar(df)
```

## Histogram of numeric variables

```{r q6}
plot_histogram(df)
```

### Data Exploration

Total Transaction Count vs. Total Transaction Amount

```{r q7}
ggplot(df, aes(Total_Trans_Amt,Total_Trans_Ct)) + geom_point()

```

# K-Means

## Recipe

```{r q8}
kmeans_recipe <- recipe(~Income_Category + Education_Level + Total_Trans_Amt + Total_Trans_Ct, data=df) %>%
   step_normalize(all_numeric_predictors()) 
```

## Model and Workflow

Model Creation

```{r q9}
kmeans_model <- k_means(num_clusters = 3) |> 
  set_engine("stats")
```

Workflow Creation and Data Fitting

```{r q10}
set.seed(11) 

kmeans_model <- k_means(num_clusters = 3) %>% 
  set_engine("stats")
  
kmeans_workflow <- workflow() %>%
    add_recipe(kmeans_recipe) %>%
    add_model(kmeans_model) %>%
    fit(data = df)

kmeans_summary <- kmeans_workflow %>% 
  extract_fit_summary()

df <- df |>
  mutate(KMeansCluster=kmeans_summary$cluster_assignments)
```

## K-Means Clusters Visualization

```{r q11}
ggplot(df, aes(x = Total_Trans_Amt, y = Total_Trans_Ct, color = KMeansCluster)) +
  geom_point() +
  labs(title = "Total Transaction Amount vs. Total Transaction Count | K-Means",
       x = "Total Transaction Amount",
       y = "Total Transaction Count") +
  theme_minimal()
```

## Summary Statistics by Cluster

```{r q12}
df |>
  group_by(KMeansCluster) |>
  summarise(count=n(), across(where(is.numeric), mean)) |>
  select(KMeansCluster, Total_Trans_Ct, Total_Trans_Amt, Total_Revolving_Bal, count)
```

## Explore metrics by cluster

```{r q13}
ggplot(df, aes(x = KMeansCluster, y = Total_Trans_Ct)) +
  geom_boxplot() +
  labs(title = "Clusters by Total Transaction Count", x = "Cluster", y = "Total Transaction Count") +
  theme_minimal()
```

```{r q14}
ggplot(df, aes(x = KMeansCluster, y = Total_Trans_Amt)) +
  geom_boxplot() +
  labs(title = "Clusters by Total Transaction Amount", x = "Cluster", y = "Total Transaction Amount") +
  theme_minimal()
```

## Cluster Distribution by Categorical Variables

```{r q15}
plot_bar(df, by = "KMeansCluster")
```

# Elbow Method

## Data Preparation

```{r q16}
mini_df <- df |> 
  select(Total_Trans_Amt, Total_Trans_Ct) |>
  mutate(across(where(is.numeric), scale))
```

```{r q17}
fviz_nbclust(mini_df, kmeans, method = "wss") # weighted sum of squares / elbow
```

# Elbow Method (Manual)

```{r q18}

max_k = 10
wss = numeric(max_k)

for(k in 1:max_k) {
  kmeans_model <- k_means(num_clusters = k) %>% 
  set_engine("stats")
  
  fit_workflow <- workflow() %>%
      add_recipe(kmeans_recipe) %>%
      add_model(kmeans_model) %>%
      fit(data = df) %>%
      extract_fit_parsnip()
  
  wss[k] <- fit_workflow$fit$tot.withinss
}
```

## Elbow Plot

```{r q19}
elbow_data <- data.frame(
  k = 1:length(wss),
  wss = wss
)

ggplot(elbow_data, aes(x = k, y = wss)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = seq(1, max_k, 1)) +
  labs(title = 'Elbow Method Plot',
       x = 'Number of clusters (k)',
       y = 'Total within-cluster sum of squares (WSS)') +
  theme_minimal()
```

## Method 2: Silhouette Method

```{r q20}
fviz_nbclust(mini_df, kmeans, method='silhouette')
```

### Save clusters to CSV file

```{r q21}
write_csv(df, "credit_card_customers_kmeans.csv")
```

```{r q22}
cluster_recipe <- recipe(~Income_Category + Education_Level + Total_Trans_Amt + Total_Trans_Ct, data=df) %>%
  step_normalize(all_numeric_predictors()) 
```

### Hierarchical Clustering

```{r q23}
hc_model <- hier_clust(linkage_method = "ward.D2") 

hc_workflow <- workflow() |>
    add_recipe(cluster_recipe) |>
    add_model(hc_model) |>
    fit(data = df)
```

## Plot the dendrogram

```{r q24}
hc_workflow |> extract_fit_engine() |> plot()
```

```{r q25}
hc_summary <- hc_workflow |> 
  extract_fit_summary(num_clusters=3)

hc_summary |> str()
```

```{r q26}
fit_hc <- hc_workflow |> extract_fit_engine()

df$HierClusters <- cutree(fit_hc, k = 3)
df$HierClusters <- factor(paste("Cluster_", df$HierClusters, sep = ""), ordered = FALSE)

```

## Plot the Hierarchical Clusters

```{r q27}
ggplot(df, aes(x = Total_Trans_Amt, y = Total_Trans_Ct, color = HierClusters)) +
  geom_point() +
  labs(title = "Total Transaction Amount Vs. Total Transaction Count",
       x = "Total Transaction Amount",
       y = "Total Transaction Count") +
  theme_minimal()
```

```{r q28}
ggplot(df, aes(x = Total_Trans_Amt, y = Total_Trans_Ct, color=Total_Relationship_Count)) + 
  geom_point() + 
  facet_wrap(~ HierClusters) + 
  labs(title = "Total Transaction Amount Vs. Total Transaction Count",
       x = "Total Transaction Amount",
       y = "Total Transaction Count") 
```

# Comparing KMeans and Hierarchical Clusters

```{r q29}
conf_mat(df, truth = KMeansCluster,
         estimate = HierClusters)
```

## Plot Hierarchical Clustering vs. K-Means

```{r q30}
ggplot(df, aes(x = Total_Trans_Amt, y = Total_Trans_Ct, color = KMeansCluster, shape=HierClusters)) +
  geom_point() +
  labs(title = "Total Transaction Amount Vs. Total Transaction Count Clustering Comparison | K-Means vs. Hierarchical Clustering",
       x = "Total Transaction Amount",
       y = "Total Transaction Count") +
  theme_minimal()
```
